---
title: "Analysis"
author: "Alexander (Sasha) Pastukhov"
date: "16 April 2019"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)

library(BayesFactor)
library(boot)
library(brms)
library(car)
library(dplyr)
library(extrafont)
library(forcats)
library(fs)
library(jmuOutlier)
library(ggplot2)
library(knitr)
library(lme4)
library(lmerTest)
library(MuMIn)
library(predictmeans)
library(readr)
library(tibble)

loadfonts(quiet = TRUE)
```


```{r Tabula rasa}
rm(list = ls())
```

```{r plotting theme, message=FALSE, warning=FALSE}
plot_theme <- 
  theme(text=element_text(family="Arial"), 
        axis.text.x = element_text(size = 8, colour = 'black'), 
        axis.text.y = element_text(size = 8, colour = 'black'), 
        axis.title.x = element_text(size = 10), 
        axis.title.y = element_text(size = 10), 
        axis.ticks = element_line(size = 0.24))
```

# Importing
CSV-file format:

* `Participant`: unique participant ID
* `Session` : session timestamp
* `Block` : block index
* `Trial` : trial index
* `OnsetDelay` : randomized stimulus onset delay in seconds.
* `Bias` : biased direction of rotation for the prime, either _"left"_ or _"right"_.
* `Shape1`, `Shape2` : shape of, respectively, _prime_ and _probe_ stimuli.
* `Response`, `Response2` : response about final direction of rotation for, respectively, _prime_ and _probe_.
* `RT1`, `RT2` : response times.


```{r Import}
read_experimental_csv <- function(filename){
  observer <- 
    read_csv(filename,
             col_types = cols(
               Participant = col_character(),
               Session = col_character(),
               Block = col_integer(),
               Trial = col_integer(),
               OnsetDelay = col_double(),
               Bias = col_character(),
               Shape1 = col_character(),
               Shape2 = col_character(),
               Response1 = col_character(),
               Response2 = col_character(),
               RT1 = col_double(),
               RT2 = col_double()))
  if (length(unique(observer$Block) > 10)){
    # first block was a training phase, needs to be dropped
    observer <- 
      observer %>%
      filter(Block != min(Block))
  }
  
  # blocks count start at 1
  observer %>%
    rename(Prime = Shape1, Probe = Shape2)
}


results <- 
  tibble(filename = as.character(dir_ls(path="Data", glob="*.csv"))) %>%
  group_by(filename) %>%
  do(read_experimental_csv(.$filename)) %>%
  mutate(Prime = factor(Prime, levels=c("heavy poles sphere", "stripes-8", "stripes-4", "stripes-2")), 
         Prime = fct_recode(Prime, Sphere = "heavy poles sphere", "Quadro band"="stripes-8", "Dual band"="stripes-4", "Single band"="stripes-2"), 
         Probe = factor(Probe, levels=c("heavy poles sphere", "stripes-8", "stripes-4", "stripes-2")), 
         Probe = fct_recode(Probe, Sphere = "heavy poles sphere", "Quadro band"="stripes-8", "Dual band"="stripes-4", "Single band"="stripes-2"),
         Survived = Response1 == Response2)
```

# Psurvival analysis

```{r Psurvival, cache=TRUE}
lag1_per_participant <-
  results %>%
  group_by(Participant, Prime, Probe) %>%
  summarise(Nsurvived = sum(Survived), 
            Ntotal = n(), 
            Psurvival = Nsurvived / Ntotal)


bootstrap_within_participant <- function(df, R=2000){
  # figuring out minimal number of trials
  # (min is just a precaution, should all be the same though)
  trial_count <-
    df %>% 
    group_by(Participant) %>%
    summarise(count = n()) %>%
    pull(count) %>%
    min(.)
    

  compute_group_Psurvival <- function(unused_data, index, df){
    df %>%
      group_by(Participant) %>%
      slice(index) %>%
      summarize(Psurvival = mean(Survived, na.rm=TRUE)) %>%
      ungroup() %>%
      summarise(Pgroup = mean(Psurvival)) %>%
      pull(Pgroup)
  }
  
  bootstrapped_P <- boot(1:trial_count, compute_group_Psurvival, R = R, df=df)
  
  bca_CIs <- boot.ci(bootstrapped_P, type="bca")
  
  data.frame(Psurvival = compute_group_Psurvival(1:trial_count, 1:trial_count, df),
             lowerCI = bca_CIs$bca[4], 
             upperCI = bca_CIs$bca[5])
}
```


Plotting dependence of Psurvival on prime and probe, analogous to earlier papers.

```{r}
set.seed(50557539)

lag1 <-
  results %>%
  group_by(Prime, Probe) %>%
  do(bootstrap_within_participant(., 200))
  

ggplot(data = lag1, aes(x = Probe, y = Psurvival, ymin=lowerCI, ymax=upperCI, color = Prime, group = Prime)) + 
  geom_errorbar(width=0.1) +
  geom_line() +
  geom_point() +
  scale_color_manual(values = c("#ed1c24", "#2ba543", "#293d9b", "#f68712")) + 
  plot_theme + 
  theme(legend.position = c(0.1, 0.2), axis.text.x = element_blank())
```

Formal statistical analysis of the dependence

```{r cache=TRUE}
set.seed(17181897)

survival <-
  results %>%
  group_by(Participant, Prime, Probe) %>%
  summarise(Nsurvived = sum(Survived), 
            Ntotal = n(), 
            Psurvival = Nsurvived/Ntotal, 
            Plogit = logit(Psurvival, adjust=0.025))

`random effects` <- lme4::glmer(cbind(Nsurvived, Ntotal) ~  (1|Participant), data= survival,  family=binomial)
`+ prime` <- update(`random effects`, .~. + Prime)
`+ probe` <- update(`+ prime`, .~. + Probe)
`+ prime:probe` <- update(`+ probe`, .~. + Prime:Probe)
anova_table <- anova(`random effects`, `+ prime`, `+ probe`, `+ prime:probe`)

models <- list(`random effects`, `+ prime`, `+ probe`, `+ prime:probe`)

anova_table$R2lr <- NA
anova_table$BF <- NA
for(imodel1 in seq(2, length(models))){
  anova_table$R2lr[imodel1] <-  as.numeric(r.squaredLR(models[[imodel1]], null=`random effects`))
  anova_table$BF[imodel1] <-exp((BIC(models[[imodel1-1]]) - BIC(models[[imodel1]]))/2)
}

kable(data.frame(anova_table), digits=c(0, 1, 1, 1, 1, 1, 0, 4, 2, 2))
```


## (Dis)similarity analysis

```{r Grouped by prime or probe shape}
lag1_Prime <-
  results %>%
  group_by(Participant, Prime) %>%
  summarise(Psurvival = mean(Survived), 
            Part = "prime") %>%
  ungroup() %>%
  rename(Shape = Prime) 


lag1_Probe <-
  results %>%
  group_by(Participant, Probe) %>%
  summarise(Psurvival = mean(Survived), 
            Part = "probe") %>%
  ungroup() %>%
  rename(Shape = Probe) 

lag1_both <-
  rbind(lag1_Prime, lag1_Probe)

cor.test(lag1_Prime$Psurvival, lag1_Probe$Psurvival)
correlationBF(lag1_Prime$Psurvival, lag1_Probe$Psurvival)
  
ggplot(data = lag1_both, aes(x = Shape, y = Psurvival, color = Shape)) + 
  geom_boxplot() +
  geom_point(position = position_dodge2(width=0.1)) + 
  scale_color_manual(values = c("#ed1c24", "#2ba543", "#293d9b", "#f68712")) + 
  facet_grid(.~Part) +
  plot_theme + 
  theme(legend.position = "none", axis.text.x = element_blank())
```


```{r Dissimilarity index}
dissimilarity <- 
  results %>%
  mutate(Dissimilarity = abs(as.integer(Prime) - as.integer(Probe))) %>%
  group_by(Participant, Dissimilarity) %>%
  summarise(Nsurvived = sum(Survived), 
            Ntotal = n(),
            Psurvival = mean(Survived), 
            Plogit = car::logit(Psurvival, , adjust=0.025))

ggplot(data = dissimilarity, aes(x = Dissimilarity, y = Psurvival)) + 
  geom_boxplot(width = 0.5, aes(group=Dissimilarity)) +
  geom_point(position = position_dodge2(width=0.1)) + 
  geom_smooth(method="lm", se=FALSE) + 
  plot_theme

`random effects` <- lme4::glmer(cbind(Nsurvived, Ntotal) ~ (1|Participant), data= dissimilarity,  family=binomial)
`+ dissimilarity` <- update(`random effects`, .~. + Dissimilarity)
anova(`random effects`, `+ dissimilarity`)

r.squaredLR(`+ dissimilarity`, null=`random effects`)

set.seed(91084893) 
dissimilarity$Participant <- as.factor(as.numeric(as.factor(dissimilarity$Participant)))
bf_random <- lmBF(Plogit ~ Participant, data=dissimilarity, whichRandom = "Participant")
bf_dissimilarity <- lmBF(Plogit ~ Participant + Dissimilarity, data=dissimilarity, whichRandom = "Participant")
extractBF(bf_dissimilarity) / extractBF(bf_random)
```

## Comparison to prior work

```{r}
persistence <- 
  survival %>%
  group_by(Prime, Probe) %>%
  summarise(Psurvival = mean(Psurvival))
```


```{r}
sensory_memory <- read_csv("Prior work/bands-sensory-memory.csv", 
                           col_types=cols(Participant = col_character(),
                                          Prime = col_character(),
                                          Probe = col_character(),
                                          Nsurvived = col_double(),
                                          Ntotal = col_double()))


sensory_memory_p <-
  sensory_memory %>%
  mutate(Psurvival = Nsurvived/Ntotal) %>%
  mutate(Prime = as.factor(Prime),
         Prime = fct_recode(Prime, "Quadro band"="Quadro", "Dual band"="Dual", "Single band"="Single"),
         Prime = fct_relevel(Prime, "Sphere", "Quadro band", "Dual band", "Single band"), 
         Probe = as.factor(Probe),
         Probe = fct_recode(Probe, "Quadro band"="Quadro", "Dual band"="Dual", "Single band"="Single"),
         Probe = fct_relevel(Probe, "Sphere", "Quadro band", "Dual band", "Single band")) %>%
  group_by(Prime, Probe) %>%
  summarise(Pmemory = mean(Psurvival))

ggplot(data=sensory_memory_p,
       aes(x = Probe, y = Pmemory, color=Prime, group=Prime)) + 
  geom_line()+
  geom_point()


persistence_memory <-
  left_join(persistence, sensory_memory_p, by=c("Prime", "Probe"))
  

set.seed(63833568)
perm.cor.test(persistence_memory$Psurvival, persistence_memory$Pmemory)
cor.test(persistence_memory$Psurvival, persistence_memory$Pmemory)
correlationBF(persistence_memory$Psurvival, persistence_memory$Pmemory)
```

```{r}
adaptation <- read_csv("Prior work/bands-adaptation.csv", 
                           col_types=cols(Participant = col_character(),
                                          Prime = col_character(),
                                          Probe = col_character(),
                                          Nsurvived = col_double(),
                                          Ntotal = col_double()))


adaptation_p <-
  adaptation %>%
  mutate(Psurvival = Nsurvived/Ntotal) %>%
   mutate(Prime = as.factor(Prime),
         Prime = fct_recode(Prime, "Quadro band"="Quadro", "Dual band"="Dual", "Single band"="Single"),
         Prime = fct_relevel(Prime, "Sphere", "Quadro band", "Dual band", "Single band"), 
         Probe = as.factor(Probe),
         Probe = fct_recode(Probe, "Quadro band"="Quadro", "Dual band"="Dual", "Single band"="Single"),
         Probe = fct_relevel(Probe, "Sphere", "Quadro band", "Dual band", "Single band")) %>%
  group_by(Prime, Probe) %>%
  summarise(Padaptation = mean(Psurvival))
  
ggplot(data = adaptation_p, 
       aes(x = Probe, y = Padaptation, color=Prime, group=Prime)) + 
  geom_line()+
  geom_point()

persistence_adaptation <- left_join(persistence, adaptation_p, by=c("Prime", "Probe"))


set.seed(45949786)
perm.cor.test(persistence_adaptation$Psurvival, persistence_adaptation$Padaptation)
cor.test(persistence_adaptation$Psurvival, persistence_adaptation$Padaptation)
correlationBF(persistence_adaptation$Psurvival, persistence_adaptation$Padaptation)
```


```{r}
history_effects <-
  rbind(persistence_memory %>%
              mutate(Effect = "sensory memory") %>%
              rename(Peffect = Pmemory),
            persistence_adaptation %>%
              mutate(Effect = "adaptation") %>%
              rename(Peffect = Padaptation))
  
  

ggplot(history_effects, aes(x = Psurvival, y=Peffect, group=Effect, color=Effect)) + 
  geom_point() + 
  geom_smooth(method = "lm") + 
  plot_theme +
  theme(legend.position = "none") + 
  ylab("Psurvival memory/adaptation") +
  xlab("Psurvival persistence") + 
  scale_y_continuous(breaks = seq(0.2, 0.9, 0.1)) + 
  ylim(0, 1)+ xlim(0.5, 1)
```
```